---
title: "Week 1 - Notes"
output:
  pdf_document: default
  html_document:
    df_print: paged
  html_notebook: default
latex_engine: xelatex
---


# Data Load

```{r global_options, include = FALSE}
knitr::opts_chunk$set(message=FALSE, tidy.opts=list(width.cutoff=60)) 
```


```{r message=FALSE, warning=FALSE, include=FALSE}
library(dplyr)
```

```{r warning=FALSE}
avgpm25_gitraw <- "https://raw.githubusercontent.com"
avgpm25_user <- "/jtleek/modules/master/04_ExploratoryAnalysis/exploratoryGraphs/data"
avgpm25_filename <- "/avgpm25.csv"
avgpm25 <- paste(avgpm25_gitraw, avgpm25_user, avgpm25_filename, sep ="")

download.file(avgpm25,destfile="./data/avgpm25.csv")

classes <- c("numeric", "character", "factor", "numeric", "numeric")

pollution <- read.csv("./data/avgpm25.csv", colClasses = classes)
head(pollution)
summary(pollution$pm25)
```


# Histograms

## Simple histogram, with:
### Vertical line at national limit of 12.
### Vertical line at median.

```{r}
hist(pollution$pm25, col = "green", breaks=200)
abline(v = 12, lwd = 2)
abline(v = median(pollution$pm25), col = "magenta", lwd = 4)
```

## Multiple

```{r}
par(mfrow = c(2, 1), mar = c(4, 4, 2, 1))
hist(subset(pollution, region == "east")$pm25, col = "green") 
hist(subset(pollution, region == "west")$pm25, col = "green")
```

# Bloxplots

## Simple

```{r}
boxplot(pollution$pm25, col = "blue") 
abline(h = 12)
```

## By region

```{r}
boxplot(pm25 ~ region, data = pollution, col = "red")
```

# Barplot

```{r}
barplot(table(pollution$region), col = "wheat", main = "Number of Counties in Each Region")
```

# Scatterplots

## Simple

```{r}
with(pollution, plot(latitude, pm25)) 
abline(h=12,lwd=2,lty=2)
```

## With color

```{r}
with(pollution, plot(latitude, pm25, col = region)) 
abline(h=12,lwd=2,lty=2)
```

## Multiple

```{r}
par(mfrow = c(1, 2), mar = c(5, 4, 2, 1))
with(subset(pollution, region == "west"), plot(latitude, pm25, main = "West")) 
with(subset(pollution, region == "east"), plot(latitude, pm25, main = "East"))
```

# Lattice

```{r}
library(lattice)
xyplot(pm25 ~ latitude | region, data = pollution)
```

# Building Plots

## Base Plotting System

```{r}
library(datasets)
data(cars)
with(cars, plot(speed, dist))
```

## Lattice System

```{r}
library(lattice)
state <- data.frame(state.x77, region = state.region) 
xyplot(Life.Exp ~ Income | region, data = state, layout = c(4, 1))
```

## ggplot2

```{r}
library(ggplot2)
data(mpg)
qplot(displ, hwy, data = mpg)
```

# Base Plots

## Base plot with annotation

```{r eval=FALSE, include=FALSE}
library(datasets)
with(airquality, plot(Wind, Ozone, main = "Ozone and Wind in New York City", type ="n"))
with(subset(airquality, Month == 5), points(Wind, Ozone, col = "blue")) with(subset(airquality, Month != 5), points(Wind, Ozone, col = "red")) legend("topright", pch = 1, col = c("blue", "red"), legend = c("May", "Other Months"))
```

## With regression line

```{r}
library(datasets)
with(airquality, plot(Wind, Ozone, main = "Ozone and Wind in New York City", pch = 20))
model <- lm(Ozone ~ Wind, airquality) 
abline(model, lwd = 2)
```


## Multiple

```{r eval=FALSE, include=FALSE}
library(datasets)
par(mfrow = c(1, 2)) with(airquality, {
plot(Wind, Ozone, main = "Ozone and Wind")
plot(Solar.R, Ozone, main = "Ozone and Solar Radiation") })
```

# Graphics Devices

where to put the plots: window, PDF, JPEG, etc

## PDF

```{r}
pdf(file = "myplot.pdf")
with(faithful, plot(eruptions, waiting))
title(main = "Old Faithful Geyser data")
dev.off()
```

## PNG

```{r}
library(datasets)
with(faithful, plot(eruptions, waiting)) 
title(main = "Old Faithful Geyser data")
dev.copy(png, file = "geyserplot.png") 
dev.off()
```

