---
title: "Week 3 - Notes"
output:
  pdf_document: default
  html_notebook: default
---

# Subsetting

```{r}
set.seed(13435)
X <- data.frame("var1"=sample(1:5),"var2"=sample(6:10),"var3"=sample(11:15))
X <- X$var2[c(1,3)] = NA
X[,1]
[1] 2 3 5 1 4
X[,var1]
[1] 2 3 5 1 4
X[(X$var1>2),]
  var1 var2 var3
2    3    7   12
3    5   NA   14
5    4    9   13
X[which(X$var2>7),]
  var1 var2 var3
4    1   10   11
5    4    9   13
X[(X$var2>7),]
     var1 var2 var3
NA     NA   NA   NA
NA.1   NA   NA   NA
4       1   10   11
5       4    9   13
```

```{r}
sort(X$var1)
[1] 1 2 3 4 5
sort(X$var2,decreasing = TRUE,na.last = TRUE)
[1] 10  9  7 NA NA
X[order(X$var1,X$var2),]
```

```{r}
arrange(X,var1)
  var1 var2 var3
1    1   10   11
2    2   NA   15
3    3    7   12
4    4    9   13
5    5   NA   14
arrange(X,desc(var1))
  var1 var2 var3
1    5   NA   14
2    4    9   13
3    3    7   12
4    2   NA   15
5    1   10   11
```

```{r}
X$var4 <- rnorm(5) 
```
add new column

```{r}
newC <- rnorm(5)
Y <- cbind(X,newC)
Y
  var1 var2 var3        var4       newC
1    2   NA   15 -1.63631251  2.5437602
2    3    7   12  0.09189581  1.5545298
3    5   NA   14 -0.64486791 -0.6192328
4    1   10   11  0.01901599 -0.9261035
5    4    9   13 -0.41504580 -0.6654995
```

```{r}
restData[restData$zipCode %in% c("21212"),]
```

```{r}
data(UCBAdmissions)
DF <- as.data.frame(UCBAdmissions)
summary(DF)
      Admit       Gender   Dept       Freq      
 Admitted:12   Male  :12   A:4   Min.   :  8.0  
 Rejected:12   Female:12   B:4   1st Qu.: 80.0  
                           C:4   Median :170.0  
                           D:4   Mean   :188.6  
                           E:4   3rd Qu.:302.5  
                           F:4   Max.   :512.0  
```
# Creating new variables

```{r}
s1 <- seq(1,10,by=2)
x <- c(1,3,8,25,100,110)
a <- seq(along=x)
a
[1] 1 2 3 4 5 6
```

# Reshaping data

```{r}
mtcars$carname <- rownames(mtcars)
carMelt <- melt(mtcars, id=c("carname", "gear", "cyl"), measure.vars = c("mpg", "hp"))
carMelt
               carname gear cyl variable value
1            Mazda RX4    4   6      mpg  21.0
2        Mazda RX4 Wag    4   6      mpg  21.0
3           Datsun 710    4   4      mpg  22.8
cylData <- dcast(carMelt, cyl ~ variable,mean)
tapply(InsectSprays$count, InsectSprays$spray, sum) 
```
- sum of count by spray value


# Package plyr

```{r}
ddply(dataframe, .(variables), method, function)
```

- for each subset of a data frame, apply function then combine results into a data frame
-     method = mutate, summarize, arrange, filter, select, etc.
-     function = how the data is going to be summarized/calculated

```{r}
a <- ddply(InsectSprays, .(spray), summarize, sum = sum(count))
```

- summarize as sum the count by spray groups

# Package dplyr - Summary

- Arrange: reorder rows 
- Filter: extract a subset of rows based on criteria
- Select: return a subset of columns
- Mutate: add variables/columns or transform variables
- Rename: rename variables
- Summarize: generate summarized statistics of different variables
- Functions:
  - 1) first argument is a data frame
  - 2) what to do with it? refer to columns without $
  - 3) result is a data frame
  
# Package dplyr - Tools

Use Week3-Chicago.R to load data into chicago

```{r}
names(chicago)
```
- Variable names
```{r}
select(chicago, city:dptp)
```
- selects all columns from city into dptp
```{r}
select(chicago, -(city:dptp))
```
- selects all columns except those from city into dptp

```{r}
subset3 <- select(chicago, ends_with("2"))
```
- Keep every variable that ends with a 2.
```{r}
chic.f <- filter(chicago, pm25tmean2 > 30)
```
- observations from the data where the pm25tmean2 value is greater than 30

```{r}
chic.f <- filter(chicago, pm25tmean2 > 30 & tmpd > 80)
select(chic.f, date, tmpd, pm25tmean2)
chicago <- arrange(chicago, date)
chicago <- rename(chicago, dewpoint = dptp, pm25 = pm25tmean2)
chicago <- mutate(chicago, pm25detrend = pm25 - mean(pm25, na.rm = TRUE)) 
```
- creates new variable that is a transformation of other 2

```{r}
chicago <- mutate(chicago, tempcat = factor(1*(tmpd>80), labels = c("cold","hot")))
```
- creates a column that indicates if tmpd is superior to 80. if yes, adds hot.
```{r}
hotcold <- group_by(chicago, tempcat)
```
- groups by tempcat
```{r}
summarize(hotcold, tmpdmean = mean(tmpd), o3max = max(o3tmean2))
```
- calculates variable by group

# Merging data

```{r}
fileUrl1 <- "https://dl.dropboxusercontent.com/u/7710864/data/reviews-apr29.csv"
fileUrl2 <- "https://dl.dropboxusercontent.com/u/7710864/data/solutions-apr29.csv"
download.file(fileUrl1,destfile="./data/reviews.csv",method="curl")
download.file(fileUrl2,destfile="./data/solutions.csv",method="curl")
reviews <- read.csv("./data/reviews.csv"); 
solutions <- read.csv("./data/solutions.csv")

```

```{r}
mergedData = merge(reviews, solution,by.x="solution_id", by.y="id",all=TRUE)

```
It merges dataframes by solution_id(reviews) and id(solutions)

# Quiz

```{r}
microdata <- "https://d396qusza40orc.cloudfront.net/getdata%2Fdata%2Fss06hid.csv"
download.file(microdata,destfile="./data/microdata.csv",method="curl")
microdata <- read.csv("./data/microdata.csv")
agricultureLogical <- (microdata$ACR == 3 & microdata$AGS == 6)
head(which(agricultureLogical))
```


```{r}
install.packages('jpeg')
library(jpeg)
fileJPEG <- "https://d396qusza40orc.cloudfront.net/getdata%2Fjeff.jpg"
download.file(fileJPEG,destfile="./data/fileJPEG.csv",method="curl")
dataJPEG <- readJPEG("./data/fileJPEG.csv",native = TRUE)
quantile(dataJPEG, c(0.3,0.8))
```


```{r}
gdp <- "https://d396qusza40orc.cloudfront.net/getdata%2Fdata%2FGDP.csv"
edu <- "https://d396qusza40orc.cloudfront.net/getdata%2Fdata%2FEDSTATS_Country.csv"
download.file(gdp,destfile="./data/gdp.csv",method="curl")
download.file(edu,destfile="./data/edu.csv",method="curl")
gdp <- read.csv("./data/gdp.csv", skip=4, dec=",")
gdp <- head(gdp,190)
edu <- read.csv("./data/edu.csv")
mergedData <- merge(gdp, edu,by.x="X", by.y="CountryCode",all=FALSE)
```


```{r}
nrow(mergedData)

mergedData[,5] <- as.numeric(gsub(",", "", mergedData[,5]))
mergedData <- mergedData[order(rank(mergedData$"X.4")),]
mergedData[13,4]
```


```{r}


mergedData[,2] <- as.numeric(mergedData[,2])
tapply(mergedData$'X.1',mergedData$Income.Group,mean)
```


```{r}
mergedData$groups <- cut(mergedData$'X.1',breaks = quantile(mergedData$'X.1',c(0,0.2,0.4,0.6,0.8,1)))
table(mergedData$groups,mergedData$Income.Group)
```



